Analysis: Machine Learning - OblAmt

### 2021-12-12 code/analysis_code/analysis-ML-OA.Rmd
Last ML model update

* Null
* Decision Tree
* Random Forest
* Bagged Tree

@@

* here: for data loading/saving
* tidyverse: for data management
* tidymodels: for data modeling
* skimr: for variable summaries
* broom.mixed: for converting bayesian models to tidy tibbles
* rpart.plot: for visualizing a decision tree
* vip: for variable importance plots
* glmnet: for lasso models
* doParallel: for parallel backend for tuning processes
* ranger: for random forest models
* baguette: for bagged decision tree
```{r libraries, include = FALSE}
knitr::opts_chunk$set(echo = TRUE)

#load required packages
library(here) #for data loading/saving
library(tidyverse) #for data management
library(tidymodels) #for data modeling
library(skimr) #for variable summaries
library(broom.mixed) #for converting bayesian models to tidy tibbles
library(rpart.plot) #for visualizing a decision tree
library(vip) #for variable importance plots
library(glmnet) #for lasso models
library(doParallel) #for parallel backend for tuning processes
library(ranger) #for random forest models
library(baguette) #for bagged decision tree

#global environment options
# formatting for script to avoid scientific notation output
options(scipen=999)
```

@@

## Load Data
Load the analysis data from the `processed_data` folder in the project file.
```{r load data}
#path to data
#note the use of the here() package and not absolute paths
data_location <- here::here("data","processed_data","analysisdata.rds")

#load data. 
analysisdata <- readRDS(data_location)

#summary of data using skimr package
skimr::skim(analysisdata)
```

@@

* Drop the unused variables
* Perform feature engineering
* Transform Incident Month into ordinal variable
* Set the random seed to 123
* Split the dataset into 70% training, 30% testing, stratified on incidentType
* 5-fold cross validation, 5 times repeated, stratified on incidentType for the CV folds
* Create a recipe for data and fitting that codes categorical variables as dummy variables
```{r data setup}
#requested amount filter
obl_data <- analysisdata %>%
              dplyr::select(state, IncidentYear, IncidentMonth, declarationType, incidentType, IncidentDuration, ResponseDays, IH, PA, HM,
                            Population, Counties, TotalAgencies, FEMACSAvg, OblAmt)

#transform incident month into ordinal variable
obl_data$IncidentMonth <- ordered(obl_data$IncidentMonth, levels = 1:12,
                                  labels = c("January", "February", "March", "April", "May", "June", "July", "August", "September", "October", "November", "December"))

#not enough values per state, so create a variable that is FEMA Regions
obl_data$FEMARegion <- ifelse((obl_data$state == "CT" |
                                 obl_data$state == "ME" |
                                 obl_data$state == "MA" |
                                 obl_data$state == "NH" |
                                 obl_data$state == "RI" |
                                 obl_data$state == "VT"),
                              "Region I",
                              ifelse((obl_data$state == "NJ" |
                                         obl_data$state == "NY" |
                                         obl_data$state == "PR" |
                                         obl_data$state == "VI"),
                                      "Region II",
                                       ifelse((obl_data$state == "DE" |
                                                 obl_data$state == "MD" |
                                                 obl_data$state == "PA" |
                                                 obl_data$state == "VA" |
                                                 obl_data$state == "DC" |
                                                 obl_data$state == "WV"),
                                              "Region III",
                                               ifelse((obl_data$state == "AL" |
                                                         obl_data$state == "FL" |
                                                         obl_data$state == "GA" |
                                                         obl_data$state == "KY" |
                                                         obl_data$state == "MS" |
                                                         obl_data$state == "NC" |
                                                         obl_data$state == "SC" |
                                                         obl_data$state == "TN"),
                                                      "Region IV",
                                                      ifelse((obl_data$state == "IL" |
                                                               obl_data$state == "IN" |
                                                               obl_data$state == "MI" |
                                                               obl_data$state == "MN" |
                                                               obl_data$state == "OH" |
                                                               obl_data$state == "WI"),
                                                            "Region V",
                                                             ifelse((obl_data$state == "AR" |
                                                                         obl_data$state == "LA" |
                                                                         obl_data$state == "NM" |
                                                                         obl_data$state == "OK" |
                                                                         obl_data$state == "TX"),
                                                                      "Region VI",
                                                                       ifelse((obl_data$state == "IA" |
                                                                                 obl_data$state == "KS" |
                                                                                 obl_data$state == "MO" |
                                                                                 obl_data$state == "NE"),
                                                                              "Region VII",
                                                                               ifelse((obl_data$state == "CO" |
                                                                                         obl_data$state == "MT" |
                                                                                         obl_data$state == "ND" |
                                                                                         obl_data$state == "SD" |
                                                                                         obl_data$state == "UT" |
                                                                                         obl_data$state == "WY"),
                                                                                      "Region VIII",
                                                                                       ifelse((obl_data$state == "AZ" |
                                                                                                 obl_data$state == "CA" |
                                                                                                 obl_data$state == "HI" |
                                                                                                 obl_data$state == "NV"),
                                                                                              "Region IX",
                                                                                               ifelse((obl_data$state == "AK" |
                                                                                                         obl_data$state == "ID" |
                                                                                                         obl_data$state == "OR" |
                                                                                                         obl_data$state == "WA"),
                                                                                                      "Region X",
                                                                                                      NA
                                                                                                      )
                                                                                              )
                                                                                      )
                                                                              )
                                                                      )
                                                            )
                                                      )
                                              )
                                     )
                              )

#drop state variable
obl_data <- obl_data %>%
              dplyr::select(-c(state))

#add to "other" category for incident Types that are infrequent
obl_data$incidentType[(obl_data$incidentType == "Chemical" |
                        obl_data$incidentType == "Coastal Storm" |
                        obl_data$incidentType == "Dam/Levee Break" |
                        obl_data$incidentType == "Mud/Landslide" |
                        obl_data$incidentType == "Volcano" |
                        obl_data$incidentType == "Terrorist" |
                        obl_data$incidentType == "Snow")] <- "Other"

@@

#save file
DT_auto_fig_file = here("results","DT-auto-plot-obl.png")
ggsave(filename = DT_auto_fig_file, plot = DT_auto_plot)

#more detailed plot
DT_detail_plot <- tree_res %>%
                    workflowsets::collect_metrics() %>%
                    dplyr::mutate(tree_depth = factor(tree_depth)) %>%
                    ggplot2::ggplot(aes(cost_complexity, mean, color = tree_depth)) +
                             geom_line(size = 1.5, alpha = 0.6) +
                             geom_point(size = 2) +
                             facet_wrap(~ .metric, scales = "free", nrow = 2) +
                             scale_x_log10(labels = scales::label_number()) +
                             scale_color_viridis_d(option = "plasma", begin = 0.9, end = 0)
DT_detail_plot

#save file
DT_detail_fig_file = here("results","DT-det-plot-obl.png")
ggsave(filename = DT_detail_fig_file, plot = DT_detail_plot)

@@

### 6. Model evaluation
```{r}
#plot the tree
rpart.plot::rpart.plot(x = workflowsets::extract_fit_parsnip(best_tree_fit)$fit,
                                         roundint = F,
                                         type = 5,
                                         digits = 5,
                                         main = "Selected Decision Tree Model")
#save file manually -- ggsave throws an error

#find predictions and intervals
tree_resid <- best_tree_fit %>%
                  broom.mixed::augment(new_data = obl_train_data) %>%
                  dplyr::select(.pred, OblAmt) %>%
                  dplyr::mutate(.resid = .pred - OblAmt)

#plot model predictions from tuned model versus actual outcomes
#geom_abline draws a 45 degree line, along which the results should fall
DT_pred_act <- ggplot2::ggplot(tree_resid, aes(x = .pred, y = OblAmt)) +
                        geom_abline(slope = 1, intercept = 0, color = "red", lty = 2) + 
                        geom_point() +
                        labs(title = "Decision Tree Fit: Predicted vs. Actual Obligated Funds",
                              x = "Predicted Obligated Funds ($)",
                              y = "Actual Obligated Funds ($)")
DT_pred_act

#save file
DT_pred_act_file = here("results","DT-pred-act-obl.png")
ggsave(filename = DT_pred_act_file, plot = DT_pred_act)

#plot model with residuals
#the geom_hline plots a straight horizontal line along which the results should fall
DT_resid <- ggplot2::ggplot(tree_resid, aes(x = as.numeric(row.names(tree_resid)), y = .resid))+
                     geom_hline(yintercept = 0, color = "red", lty = 2) +
                     geom_point() +
                     labs(title = "Decision Tree Fit: Residuals",
                          x = "Observation Number",
                          y = "Residual")
DT_resid

#save file
DT_resid_file = here("results","DT-resid-obl.png")
ggsave(filename = DT_resid_file, plot = DT_resid)

#plot model fit vs residuals
#the geom_hline plots a straight horizontal line along which the results fall
DT_pred_resid <- ggplot2::ggplot(tree_resid, aes(x = .pred, y = .resid))+
                          geom_hline(yintercept = 0, color = "red", lty = 2) +
                          geom_point() +
                          labs(title = "Decision Tree Fit: Residuals vs Fitted Obligated Funds",
                                x = "Predicted Obligated Funds ($)",
                                y = "Residual")
DT_pred_resid

#save file
DT_pred_resid_file = here("results","DT-pred-resid-obl.png")
ggsave(filename = DT_pred_resid_file, plot = DT_pred_resid)

@@

@@

#save file
RF_auto_fig_file = here("results","RF-auto-plot-obl.png")
ggsave(filename = RF_auto_fig_file, plot = RF_auto_plot)
```

@@

#plot most important predictors in the model
RF_plot <- vip::vip(x_RF, num_features = 20)
RF_plot

#save  manually - ggsave throws an error

#find predictions and intervals
RF_resid <- best_RF_fit %>%
                broom.mixed::augment(new_data = obl_train_data) %>%
                dplyr::select(.pred, OblAmt) %>%
                dplyr::mutate(.resid = .pred - OblAmt)

#plot model predictions from tuned model versus actual outcomes
#geom_abline is a 45 degree line, along which the results should fall
RF_pred_act <- ggplot2::ggplot(RF_resid, aes(x = .pred, y = OblAmt)) +
                        geom_abline(slope = 1, intercept = 0, color = "red", lty = 2) +
                        geom_point() +
                        labs(title = "RF Fit: Actual vs. Predicted Obligated Funds",
                              x = "Predicted Obligated Funds ($)",
                              y = "Actual Obligated Funds ($)")

#save file
RF_pred_act_file = here("results","RF-pred-act-obl.png")
ggsave(filename = RF_pred_act_file, plot = RF_pred_act)

#plot model with residuals
#the geom_hline plots a straight horizontal line along which the results should fall
RF_resids <- ggplot2::ggplot(RF_resid, aes(x = as.numeric(row.names(RF_resid)), y = .resid))+
                            geom_hline(yintercept = 0, color = "red", lty = 2) +
                            geom_point() +
                            labs(title = "RF Fit: Residuals",
                                  x = "Observation Number",
                                  y = "Residual")
RF_resids

#save file
RF_resid_file = here("results","RF-resid-obl.png")
ggsave(filename = RF_resid_file, plot = RF_resids)

#plot model fit vs residuals
#the geom_hline plots a straight horizontal line along which the results fall
RF_pred_resid <- ggplot2::ggplot(RF_resid, aes(x = .pred, y = .resid))+
                          geom_hline(yintercept = 0, color = "red", lty = 2) +
                          geom_point() +
                          labs(title = "Random Forest Fit: Residuals vs Fitted Obligated Funds",
                                x = "Fitted Requested Funds ($)",
                                y = "Residual")
RF_pred_resid

#save file
RF_pred_resid_file = here("results","RF-pred-resid-obl.png")
ggsave(filename = RF_pred_resid_file, plot = RF_pred_resid)

@@

@@

### 5. Identify Best Model
```{r}
#select the BT model with the lowest rmse
BT_lowest_rmse <- BT_res %>%
                        tune::select_best("rmse")

#finalize the workflow by using the selected BT model
best_BT_wflow <- BT_wflow %>%
                      tune::finalize_workflow(BT_lowest_rmse)
best_BT_wflow

#one last fit on the training data
best_BT_fit <- best_BT_wflow %>%
                    parsnip::fit(data = obl_train_data)
```

@@

### 6. Model evaluation
```{r}
#extract model from final fit
x_BT <- best_BT_fit$fit$fit$fit

#vip package doesn't support BT

#find predictions and intervals
BT_resid <- best_BT_fit %>%
                  broom.mixed::augment(new_data = obl_train_data) %>%
                  dplyr::select(.pred, OblAmt) %>%
                  dplyr::mutate(.resid = .pred - OblAmt)

#plot model predictions from tuned model versus actual outcomes
#geom_abline draws a 45 degree line, along which the results should fall
BT_pred_act <- ggplot2::ggplot(BT_resid, aes(x = .pred, y = OblAmt)) +
                        geom_abline(slope = 1, intercept = 0, color = "red", lty = 2) + 
                        geom_point() +
                        labs(title = "Bagged Tree Fit: Predicted vs. Actual Obligated Funds",
                              x = "Predicted Obligated Funds ($)",
                              y = "Actual Obligated Funds ($)")
BT_pred_act

#save file
BT_pred_act_file = here("results","BT-pred-act-obl.png")
ggsave(filename = BT_pred_act_file, plot = BT_pred_act)

#plot model with residuals
#the geom_hline plots a straight horizontal line along which the results should fall
BT_resids <- ggplot2::ggplot(BT_resid, aes(x = as.numeric(row.names(BT_resid)), y = .resid))+
                      geom_hline(yintercept = 0, color = "red", lty = 2) +
                      geom_point() +
                      labs(title = "Bagged Tree Fit: Residuals",
                            x = "Observation Number",
                            y = "Residual")
BT_resids

#save file
BT_resid_file = here("results","BT-resids-obl.png")
ggsave(filename = BT_resid_file, plot = BT_resids)

#plot model fit vs residuals
#the geom_hline plots a straight horizontal line along which the results fall
BT_pred_resid <- ggplot2::ggplot(BT_resid, aes(x = .pred, y = .resid))+
                          geom_hline(yintercept = 0, color = "red", lty = 2) +
                          geom_point() +
                          labs(title = "Bagged Tree Fit: Residuals vs Fitted Obligated Funds",
                                x = "Predicted Obligated Funds ($)",
                                y = "Residual")
BT_pred_resid

#save file
BT_pred_resid_file = here("results","BT-pred-resid-obl.png")
ggsave(filename = BT_pred_resid_file, plot = BT_pred_resid)

#print model performance
#print 10 best performing hyperparameter sets
BT_res %>%
  tune::show_best(n = 10) %>%
  dplyr::select(rmse = mean, std_err, cost_complexity) %>%
  dplyr::mutate(rmse = round(rmse, 3),
                std_err = round(std_err, 4),
                cost_complexity = scales::scientific(cost_complexity))

#print the best model performance
BT_performance <- BT_res %>% tune::show_best(n = 1)
print(BT_performance)

#compare model performance to null model
BT_RMSE <- BT_res %>%
                tune::show_best(n = 1) %>%
                dplyr::transmute(
                  rmse = round(mean, 3),
                  SE = round(std_err, 4),
                  model = "Bagged Tree") %>%
               dplyr::bind_rows(RF_RMSE)
BT_RMSE

#save combined table
model_comp_file = here("results", "ML-model-comp-obl.Rds")
saveRDS(BT_RMSE, file = model_comp_file)
```

In examining the models, the Bagged Tree has the best RMSE, but it also has a higher SE than the other models. We will still choose the bagged tree as our model on which to evaluate the test data.

--

## Model Selection and Evaluation
Using the bagged tree model
```{r}
#fit model to training set but evaluate on the test data
BT_fit_test <- best_BT_wflow %>%
                    tune::last_fit(split = obl_data_split)

#compare test performance against training performance
BT_rmse_test <- collect_metrics(BT_fit_test) %>%
  dplyr::select(rmse = .estimate) %>%
  dplyr::mutate(data = "test")

BT_comp <- BT_RMSE %>%
                  dplyr::filter(model == "Bagged Tree") %>%
                  dplyr::transmute(rmse, data = "train") %>%
                  dplyr::bind_rows(BT_rmse_test) %>%
                  dplyr::slice(-3) #don't know why the third row shows up
BT_comp
#RMSEs are incredibly different...

#find predictions and intervals
BT_resid_fit <- BT_fit_test %>%
                  broom.mixed::augment() %>%
                  dplyr::select(.pred, OblAmt) %>%
                  dplyr::mutate(.resid = .pred - OblAmt)

#plot model predictions from tuned model versus actual outcomes
#geom_abline draws a 45 degree line, along which the results should fall
BT_pred_act_test <- ggplot2::ggplot(BT_resid_fit, aes(x = .pred, y = OblAmt)) +
                        geom_abline(slope = 1, intercept = 0, color = "red", lty = 2) + 
                        geom_point() +
                        labs(title = "Bagged Tree Test Fit: Predicted vs. Actual Obligated Funds",
                              x = "Predicted Obligated Funds ($)",
                              y = "Actual Obligated Funds ($)")
BT_pred_act_test

#save file
BT_pred_act_test_file = here("results","BT-pred-act-test-obl.png")
ggsave(filename = BT_pred_act_test_file, plot = BT_pred_act_test)

#plot model with residuals
#the geom_hline plots a straight horizontal line along which the results should fall
BT_resids_test <- ggplot2::ggplot(BT_resid_fit, aes(x = as.numeric(row.names(BT_resid_fit)), y = .resid))+
                      geom_hline(yintercept = 0, color = "red", lty = 2) +
                      geom_point() +
                      labs(title = "Bagged Tree Test Fit: Residuals",
                            x = "Observation Number",
                            y = "Residual")
BT_resids_test

#save file
BT_resid_test_file = here("results","BT-resid-test-obl.png")
ggsave(filename = BT_resid_test_file, plot = BT_resids_test)

#plot model fit vs residuals
#the geom_hline plots a straight horizontal line along which the results fall
BT_pred_resid_test <- ggplot2::ggplot(BT_resid_fit, aes(x = .pred, y = .resid))+
                          geom_hline(yintercept = 0, color = "red", lty = 2) +
                          geom_point() +
                          labs(title = "Bagged Tree Test Fit: Residuals vs Fitted Obligated Funds",
                                x = "Predicted Obligated Funds ($)",
                                y = "Residual")
BT_pred_resid_test

#save file
BT_pred_resid_test_file = here("results","BT-pred-resid-test-obl.png")
ggsave(filename = BT_pred_resid_test_file, plot = BT_pred_resid_test)
```

>>>>

* Null
* Decision Tree
* Random Forest
* Elastic Net
* LASSO

@@

* here: for data loading/saving
* tidyverse: for data management
* tidymodels: for data modeling
* skimr: for variable summaries
* spatialsample: for spatial cross-validation
* broom.mixed: for converting bayesian models to tidy tibbles
* rpart.plot: for visualizing a decision tree
* vip: for variable importance plots
* glmnet: for lasso models
* doParallel: for parallel backend for tuning processes
* ranger: for random forest models
* viridis: for a consistent color scheme in figures
```{r libraries, include = FALSE}
knitr::opts_chunk$set(echo = TRUE)

#load required packages
library(here) #for data loading/saving
library(tidyverse) #for data management
library(tidymodels) #for data modeling
library(skimr) #for variable summaries
library(spatialsample) #for spatial cross validation
library(broom.mixed) #for converting bayesian models to tidy tibbles
library(rpart.plot) #for visualizing a decision tree
library(vip) #for variable importance plots
library(glmnet) #for lasso models
library(doParallel) #for parallel backend for tuning processes
library(ranger) #for random forest models
library(viridis) #for a consistent color scheme in figures

#global environment options
# formatting for script to avoid scientific notation output
options(scipen=999)

#set ggplot theme to classic
ggplot2::theme_set(theme_classic())

#center plot titles
theme_update(plot.title = element_text(hjust = 0.5))
```

@@

## Load Data
Load the OblAmt analysis data from the `processed_data` folder in the project file.
```{r load data}
#path to data
#note the use of the here() package and not absolute paths
data_location <- here::here("data","processed_data","analysisdata-OA.rds")

#load data. 
obl_data <- readRDS(data_location)

#summary of data using skimr package
skimr::skim(obl_data)
```

@@

* Log transform the outcome
* Drop `ReqAmt`, `disasterNumber`, `declarationTitle`
* Set the random seed to 123
* Split the dataset into 70% training, 30% testing
* 5-fold cross validation for the spatial CV folds
* Create a recipe for data and fitting that codes categorical variables as dummy variables
```{r data setup}
#log transform ReqAmt
obl_data$logOblAmt <- log(obl_data$OblAmt)

#drop unnecessary variables
obl_data <- obl_data %>%
              dplyr::select(-c(OblAmt, disasterNumber, declarationTitle))

#specify all character variables as factors and clean variable names
obl_data <- obl_data %>% 
              janitor::clean_names() %>%
              dplyr::mutate_if(is.character, factor)

@@

#save file
DT_auto_fig_file = here("results","DT-auto-plot-OA.png")
ggsave(filename = DT_auto_fig_file, plot = DT_auto_plot)

#more detailed plot
DT_detail_plot <- tree_res %>%
                    workflowsets::collect_metrics() %>%
                    dplyr::mutate(tree_depth = factor(tree_depth)) %>%
                    ggplot2::ggplot(aes(cost_complexity, mean, color = tree_depth)) +
                             geom_line(size = 1.5, alpha = 0.6) +
                             geom_point(size = 2) +
                             facet_wrap(~ .metric, scales = "free") +
                             scale_x_log10(labels = scales::label_number()) +
                             scale_color_viridis_d(option = "plasma", begin = 0.9, end = 0) +
                             labs(x = "cost complexity",
                                  y = "mean",
                                  color = "Tree Depth")
DT_detail_plot

#save file
DT_det_fig_file = here("results","DT-det-plot-OA.png")
ggsave(filename = DT_det_fig_file, plot = DT_detail_plot)

@@

### 6. Model evaluation
```{r}
#plot the tree
rpart.plot::rpart.plot(x = workflowsets::extract_fit_parsnip(best_tree_fit)$fit,
                                                     roundint = F,
                                                     type = 5,
                                                     digits = 5,
                                                     main = "Selected Decision Tree Model",
                                                     box.palette = viridis::viridis(10, option = "D", begin = 0.9, end = 0),
                                                     shadow.col = "grey65",
                                                     col = "grey99")
#save manually as "DT-plot-OA.png" -- ggsave throws an error

#find predictions and intervals
tree_resid <- best_tree_fit %>%
                  broom.mixed::augment(new_data = obl_train_data) %>%
                  dplyr::select(.pred, log_obl_amt) %>%
                  dplyr::mutate(.resid = .pred - log_obl_amt)

#plot model predictions from tuned model versus actual outcomes
#geom_abline draws a 45 degree line, along which the results should fall
DT_pred_act <- ggplot2::ggplot(tree_resid, aes(x = .pred, y = log_obl_amt)) +
                        geom_abline(slope = 1, intercept = 0, color = "red", lty = 2) + 
                        geom_point(color = "#3b528b") +
                        scale_y_continuous(expand = c(0,0),
                                           limits = c(-30, 25)) +
                        scale_x_continuous(limits = c(0, 25)) +
                        labs(title = "Decision Tree Fit: Predicted vs. Actual Log(Obligated Funds)",
                              x = "Predicted Log(Obligated Funds) ($)",
                              y = "Actual Log(Obligated Funds) ($)") 
DT_pred_act

#save file
DT_pred_act_file = here("results","DT-pred-act-OA.png")
ggsave(filename = DT_pred_act_file, plot = DT_pred_act)

#plot model with residuals
#the geom_hline plots a straight horizontal line along which the results should fall
DT_resid <- ggplot2::ggplot(tree_resid, aes(x = as.numeric(row.names(tree_resid)), y = .resid))+
                     geom_hline(yintercept = 0, color = "red", lty = 2) +
                     geom_point(color = "#3b528b") +
                     labs(title = "Decision Tree Fit: Residuals",
                          x = "Observation Number",
                          y = "Residual")
DT_resid

#save file
DT_resid_file = here("results","DT-resid-OA.png")
ggsave(filename = DT_resid_file, plot = DT_resid)

#plot model fit vs residuals
#the geom_hline plots a straight horizontal line along which the results fall
DT_pred_resid <- ggplot2::ggplot(tree_resid, aes(x = .pred, y = .resid))+
                          geom_hline(yintercept = 0, color = "red", lty = 2) +
                          geom_point(color = "#3b528b") +
                          labs(title = "Decision Tree Fit: Residuals vs Fitted Log(Obligated Funds)",
                                x = "Predicted Log(Obligated Funds) ($)",
                                y = "Residual")
DT_pred_resid

#save file
DT_pred_resid_file = here("results","DT-pred-resid-OA.png")
ggsave(filename = DT_pred_resid_file, plot = DT_pred_resid)

@@

The identified best performing tree model has some interesting results. It identifies percent of state counties awarded IH program, New York, longitude, total number of agencies, FEMA cost share average, total response days, and an incident month of November for predicting requested FEMA funding. New York and November are particularly surprising key predictors. It is interesting that incident type or declaration type are not included as significant. The RMSE actually increases from the Null, which suggests this is not an appropriate model. Moving on.

@@

#save file
RF_auto_fig_file = here("results","RF-auto-plot-OA.png")
ggsave(filename = RF_auto_fig_file, plot = RF_auto_plot)
```

@@

#plot most important predictors in the model
RF_plot <- vip::vip(x_RF, 
                    num_features = 20,
                    aesthetics = list(fill = viridis::viridis(20, option = "D", begin = 0.9, end = 0)))
RF_plot

#save manually as "RF-plot-OA.png" -- ggsave throws errors

#find predictions and intervals
RF_resid <- best_RF_fit %>%
                broom.mixed::augment(new_data = obl_train_data) %>%
                dplyr::select(.pred, log_obl_amt) %>%
                dplyr::mutate(.resid = .pred - log_obl_amt)

#plot model predictions from tuned model versus actual outcomes
#geom_abline is a 45 degree line, along which the results should fall
RF_pred_act <- ggplot2::ggplot(RF_resid, aes(x = .pred, y = log_obl_amt)) +
                        geom_abline(slope = 1, intercept = 0, color = "red", lty = 2) +
                        geom_point(color = "#3b528b") +
                        scale_y_continuous(expand = c(0,0),
                                           limits = c(-30, 25)) +
                        scale_x_continuous(limits = c(0, 25)) +
                        labs(title = "RF Fit: Actual vs. Predicted Log(Obligated Funds)",
                              x = "Predicted Log(Obligated Funds) ($)",
                              y = "Actual Log(Obligated Funds) ($)")
RF_pred_act

#save file
RF_pred_act_file = here("results","RF-pred-act-OA.png")
ggsave(filename = RF_pred_act_file, plot = RF_pred_act)

#plot model with residuals
#the geom_hline plots a straight horizontal line along which the results should fall
RF_resids <- ggplot2::ggplot(RF_resid, aes(x = as.numeric(row.names(RF_resid)), y = .resid))+
                            geom_hline(yintercept = 0, color = "red", lty = 2) +
                            geom_point(color = "#3b528b") +
                            labs(title = "RF Fit: Residuals",
                                  x = "Observation Number",
                                  y = "Residual")
RF_resids

#save file
RF_resid_file = here("results","RF-resid-OA.png")
ggsave(filename = RF_resid_file, plot = RF_resids)

#plot model fit vs residuals
#the geom_hline plots a straight horizontal line along which the results fall
RF_pred_resid <- ggplot2::ggplot(RF_resid, aes(x = .pred, y = .resid))+
                          geom_hline(yintercept = 0, color = "red", lty = 2) +
                          geom_point(color = "#3b528b") +
                          labs(title = "Random Forest Fit: Residuals vs Fitted Log(Obligated Funds)",
                                x = "Fitted Log(Obligated Funds) ($)",
                                y = "Residual")
RF_pred_resid

#save file
RF_pred_resid_file = here("results","RF-pred-resid-OA.png")
ggsave(filename = RF_pred_resid_file, plot = RF_pred_resid)

@@

In examining the results of the RF model within the context of RMSE, it performs marginally better than the decision tree. It identifies total number of federal agencies involved, percent of state counties awarded IH, FEMA cost share average, percent of counties awarded HM, and incident duration as the five most important predictors. Let's try another model to see if we can find an improvement in the RMSE.

@@

### 4. Tuning Using Cross-Validation and the `tune_grid()` function
```{r}
#tune the model with previously specified cross-validation and RMSE as target metric
EN_res <- elastic_wflow %>%
              tune::tune_grid(resamples = obl_folds_spatial,
                              grid = EN_grid,
                              control = control_grid(verbose = TRUE, save_pred = TRUE),
                              metrics = metric_set(rmse))

#look at top 5 EN models
top_EN_models <- EN_res %>%
                    tune::show_best("rmse", n = 5)
top_EN_models

#default visualization
EN_auto_plot <- EN_res %>% autoplot()
EN_auto_plot

#save file
EN_auto_fig_file = here("results","EN-auto-plot-OA.png")
ggsave(filename = EN_auto_fig_file, plot = EN_auto_plot)
```

@@

### 5. Identify Best Model
```{r}
#select the EN model with the lowest rmse
EN_lowest_rmse <- EN_res %>%
                      tune::select_best("rmse")

#finalize the workflow by using the selected RF model
best_EN_wflow <- elastic_wflow %>%
                      tune::finalize_workflow(EN_lowest_rmse)
best_EN_wflow

#one last fit on the training data
best_EN_fit <- best_EN_wflow %>%
                    parsnip::fit(data = obl_train_data)
```

### 6. Model evaluation
```{r}
#extract model from final fit
x_EN <- best_EN_fit$fit$fit$fit

#plot most important predictors in the model
EN_plot <- vip::vip(x_EN, 
                    num_features = 20,
                    aesthetics = list(fill = viridis::viridis(20, option = "D", begin = 0.9, end = 0)))
EN_plot

#save manually as "EN-plot-OA.png" -- ggsave throws errors

#find predictions and intervals
EN_resid <- best_EN_fit %>%
                broom.mixed::augment(new_data = obl_train_data) %>%
                dplyr::select(.pred, log_obl_amt) %>%
                dplyr::mutate(.resid = .pred - log_obl_amt)

#plot model predictions from tuned model versus actual outcomes
#geom_abline is a 45 degree line, along which the results should fall
EN_pred_act <- ggplot2::ggplot(EN_resid, aes(x = .pred, y = log_obl_amt)) +
                        geom_abline(slope = 1, intercept = 0, color = "red", lty = 2) +
                        geom_point(color = "#3b528b") +
                        scale_y_continuous(expand = c(0,0),
                                           limits = c(-30, 25)) +
                        scale_x_continuous(limits = c(0, 25)) +
                        labs(title = "EN Fit: Actual vs. Predicted Log(Obligated Funds)",
                              x = "Predicted Log(Obligated Funds) ($)",
                              y = "Actual Log(Obligated Funds) ($)")
EN_pred_act

#save file
EN_pred_act_file = here("results","EN-pred-act-OA.png")
ggsave(filename = EN_pred_act_file, plot = EN_pred_act)

#plot model with residuals
#the geom_hline plots a straight horizontal line along which the results should fall
EN_resids <- ggplot2::ggplot(EN_resid, aes(x = as.numeric(row.names(EN_resid)), y = .resid))+
                            geom_hline(yintercept = 0, color = "red", lty = 2) +
                            geom_point(color = "#3b528b") +
                            labs(title = "EN Fit: Residuals",
                                  x = "Observation Number",
                                  y = "Residual")
EN_resids

#save file
EN_resid_file = here("results","EN-resid-OA.png")
ggsave(filename = EN_resid_file, plot = EN_resids)

#plot model fit vs residuals
#the geom_hline plots a straight horizontal line along which the results fall
EN_pred_resid <- ggplot2::ggplot(EN_resid, aes(x = .pred, y = .resid))+
                          geom_hline(yintercept = 0, color = "red", lty = 2) +
                          geom_point(color = "#3b528b") +
                          labs(title = "EN Fit: Residuals vs Fitted Log(Obligated Funds)",
                                x = "Fitted Log(Obligated Funds) ($)",
                                y = "Residual")
EN_pred_resid

#save file
EN_pred_resid_file = here("results","EN-pred-resid-OA.png")
ggsave(filename = EN_pred_resid_file, plot = EN_pred_resid)

#print the 10 best performing hyperparameter sets
EN_res %>%
  tune::show_best(n = 10) %>%
  dplyr::select(rmse = mean, std_err)

#print the best model
EN_performance <- EN_res %>% tune::show_best(n = 1)
EN_performance

#compare model performance to null model (and other models)
EN_RMSE <- EN_res %>%
              tune::show_best(n = 1) %>%
              dplyr::transmute(
                rmse = round(mean, 3),
                SE = round(std_err, 4),
                model = "EN") %>%
             dplyr::bind_rows(RF_RMSE)
EN_RMSE
```
The Elastic Net model places great emphasis on FEMA Cost Share average above all other predictors. The next closest predictors include Hawaii, New York, New Mexico, and DC as key predictors. The EN model has an improved RMSE, but the null model technically still has a lower RMSE.

## LASSO Model

#define the LASSO model
#mixture = 1 identifies the model to be a LASSO model
lasso_mod <- linear_reg(mode = "regression",
                           penalty = tune(),
                           mixture = 1) %>%
                parsnip::set_engine("glmnet")
```

### 2. Workflow Definition
```{r}
#define lasso workflow
lasso_wflow <- workflows::workflow() %>%
                 workflows::add_model(lasso_mod) %>%
                 workflows::add_recipe(OA_rec)
```

### 3. Tuning Grid Specification
```{r}
#tuning grid specification
lasso_grid <- tibble(penalty = 10^seq(-3, 0, length.out = 30))

#5 lowest penalty values
lasso_grid %>%
  dplyr::top_n(-5)

#5 highest penalty values
lasso_grid %>%
  dplyr::top_n(5)
```

### 4. Tuning Using Cross-Validation and the `tune_grid()` function
```{r}
#tune the model with previously specified cross-validation and RMSE as target metric
lasso_res <- lasso_wflow %>%
              tune::tune_grid(resamples = obl_folds_spatial,
                              grid = lasso_grid,
                              control = control_grid(verbose = TRUE, save_pred = TRUE),
                              metrics = metric_set(rmse))

#look at top 5 EN models
top_lasso_models <- lasso_res %>%
                        tune::show_best("rmse", n = 5)
top_lasso_models

#default visualization
lasso_auto_plot <- lasso_res %>% autoplot()
lasso_auto_plot

#save file
lasso_auto_fig_file = here("results","lasso-auto-plot-OA.png")
ggsave(filename = lasso_auto_fig_file, plot = lasso_auto_plot)
```

### 5. Identify Best Model
```{r}
#select the LASSO model with the lowest rmse
lasso_lowest_rmse <- lasso_res %>%
                        tune::select_best("rmse")

#finalize the workflow by using the selected lasso model
best_lasso_wflow <- lasso_wflow %>%
                      tune::finalize_workflow(lasso_lowest_rmse)
best_lasso_wflow

#one last fit on the training data
best_lasso_fit <- best_lasso_wflow %>%
                    parsnip::fit(data = obl_train_data)
```

### 6. Model evaluation
```{r}
#extract model from final fit
x_lasso <- best_lasso_fit$fit$fit$fit

#plot most important predictors in the model
lasso_plot <- vip::vip(x_lasso, 
                    num_features = 20,
                    aesthetics = list(fill = viridis::viridis(20, option = "D", begin = 0.9, end = 0)))
lasso_plot

#save manually as "LASSO-plot-OA.png" -- ggsave throws errors

#find predictions and intervals
lasso_resid <- best_lasso_fit %>%
                broom.mixed::augment(new_data = obl_train_data) %>%
                dplyr::select(.pred, log_obl_amt) %>%
                dplyr::mutate(.resid = .pred - log_obl_amt)
